For submission to the Journal of Integrated Pest Management

### 2021-04-09 ofv_fife_2021.Rmd
fixed error, added html
Hacked down the abstract to fit under 250 word limit. Probably will need to be edited again, sent it to other authors

Several flat mite species, all from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae), are the only known vectors for the plant viruses known as cileviruses and dichorhaviruses. The *B. californicus* species group exclusively can transmit Orchid fleck dichorhavirus (OFV) in a persistent propagative manner.
OFV is the type species for the genus *Dichorhavirus* and infects more than fifty plant species belonging to the family Orchidaceae, Asparagaceae (Nolinoidaea), and Rutaceae (*Citrus*). During the summer of 2020, chlorotic ringspot symptoms were seen on Giant Lilyturf (*Liriope* spp., cv. ‘Gigantea’, Asparagaceae: Nolinoidaea) and leaf chlorosis was observed on leaves of *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in a landscape in Leon County, Florida. In both cases, the presence of OFV was confirmed using OFV specific conventional reverse transcription polymerase chain assay (RT-PCR) assay and Sanger sequencing. RT-PCR amplicons had a 98\% identity with the known OFV sequences available in the GenBank (Accession numbers: AB244418 and LC222630). The identification was also confirmed with the quantitative RT-PCR (RT-qPCR). Additional leaf samples were collected from other Nolinoidaea, including Aztec Grass (*Ophiopogon intermedius* cv. 'Argenteomarginatus' D. Don, Asparagaceae: Nolinoidaea), Mondo Grass *O. japonicus* and Lilyturf *L. muscari*, with ringspot symptoms were observed in Leon and Alachua counties. Identification of partial genome sequence confirmed the presence of both the orchid strains (OFV-Orc1 and OFV-Orc2) in Florida. Three mite species were identified from OFV-infected plants: *Brevipalpus californicus* sensu lato, *B. obovatus* Donnadieu and *B. confusus* Banks. At least one of these species is presumably responsible for OFV transmission. Florida has various *Brevipalpus*susceptible native and introduced plant species in the landscape. In this study we are reporting three new hosts from the family Asperagaceae from multiple locations. Our data suggested that the OFV is widely distributed in Florida and might be a potential threat for *Liriope* spp., *Ophiopogon* spp. and *Aspidistra elatior* which are commonly used in landscaping in the southeastern United States, if not controlled. As both the orchid strains of OFV are known to infect citrus and cause citrus leprosis disease, a survey in citrus growing regions of Florida is essential, emphasizing plants within the families Orchidaceae, Rutaceae and Asparagaceae.

>>>>

Orchid Fleck Virus (OFV) is the type species for the genus *Dichorhavirus*, a plant pathogen which infects more than fifty plant species belonging to the family Orchidaceae, Asparagaceae (Nolinoidaea), and Rutaceae (*Citrus*). The sole vectors for dichorhaviruses are mites from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae), which are also known to spread cileviruses. OFV was found infecting liriopogons *Liriope* spp. & *Ophiopogon* spp. (Asparagaceae: Nolinoidaea) and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea), during the summer of 2020 in a landscape in Leon County, Florida. OFV specific conventional reverse transcription polymerase chain assay (RT-PCR) assay and Sanger sequencing. RT-PCR amplicons had a 98\% identity with the known OFV sequences available in the GenBank (Accession numbers: AB244418 and LC222630). Virus identity was also confirmed via quantitative RT-PCR (RT-qPCR). Additional leaf samples were collected from other putatively infected Asparagaceae from Leon and Alachua counties. Identification of partial genome sequence confirmed the presence of both OFV orchid strains (OFV-Orc1 and OFV-Orc2) in Florida. These strains of OFV are known to infect citrus and cause citrus leprosis disease. Three potential vector species were identified from OFV-infected plants: *Brevipalpus californicus* sensu lato, *B. obovatus* Donnadieu and *B. confusus* Banks. Florida has various *Brevipalpus* susceptible native and introduced plant species in the landscape. In this study we report three new hosts from the family Asperagaceae from multiple locations. Our data suggested that the OFV widely distributed in Florida and might be a threat for various plant species and possibly citrus growing in the southeastern US.
